---
title: "Machine learning models using workflowsets to classify unknown metabolites as lipids or not"
author: "Christelle Colin-Leitzinger"
date: '`r Sys.Date()`'
output: 
  html_document:
    toc: true
    toc_float: true
    toc_collapsed: false
    theme: united
    highlight: pygments
    df_print: paged
editor_options: 
  chunk_output_type: console
---

<style type="text/css">
.figure {
   margin-top: 25px;
   margin-bottom: 10px;
}

table {
    margin-top: 10px;
    margin-bottom: 25px;
}
</style>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,
                      warning = FALSE,
                      message = FALSE,
                      cache = FALSE,
                      fig.align='center'
                      )
```


```{r library}
library(tidyverse)
library(ComplexHeatmap)
library(corrplot)
library(ggcorrplot)
library(tidymodels)
library(themis)
library(discrim)
library(klaR)
library(rpart.plot)
library(vip) 
```


```{r load}
# Samples, retention time and m/z data
omics_data <- read_delim(paste0(here::here(), "/new_process_iron_log2_merged.txt")) %>% 
  janitor::clean_names()
#
metabolites_data <- read_delim(paste0(here::here(), "/hmdb_keep_v4_python_blessed.txt.zip")) %>% 
  janitor::clean_names()
```

```{r merge clean data}
unknown_metabolites <- omics_data %>% 
  # Filter identified metabolites
  filter(non_heavy_identified_flag != 1) %>% 
  dplyr::select(hmdb, row_id, 
         row_m_z, row_retention_time,
         starts_with("t"))

known_metabolites <- omics_data %>% 
  # Filter identified metabolites
  filter(non_heavy_identified_flag == 1 & !is.na(hmdb) & !str_detect(hmdb, "\\|")) %>% 
  # cleaning retention time to eliminate false detection during set up, washing and equilibrating the HPLC
  filter(row_retention_time > 0.5 & row_retention_time < 14)

mldata <- known_metabolites %>% 
  # Join with known metabolites
  inner_join(., metabolites_data %>% 
              dplyr::select(accession, taxonomy_super_class),
            by = c("hmdb" = "accession")) %>% 
  filter(!is.na(taxonomy_super_class)) %>% 
  mutate(is_lipids = case_when(
    str_detect(taxonomy_super_class, "Lipids")              ~ "Yes",
    TRUE                                                    ~ "No"
  ))
```


# **In the html, you can find the prediction of unknown metabolites using Samples, retention time and m/z information as predictors in part I and using retention time and m/z information alone as predictors in part II.**
We are trying to predict metabolite classifications. In our proteomic analysis, about 88% of the metabolites are not identified. We want to see if we can predict with a good accuracy the classification of these metabolites based on **retention time and m/z** information (Part II). We also want to assess if using **patients samples as predictors** in addition to theses 2 parameters can increase the prediction (Part I).  
<br>

***
<br>

# I. Prediction using Samples, retention time and m/z information as predictors.


## 1. Data preprocessing

- We are focusing on predicting if a metabolite is a **lipid**, **Yes/No**. I am using a variable I created, `is_lipids`, as the outcome.

- Patient samples present NO missing values :

### Missing data
```{r}
library(finalfit) 

known_metabolites %>%
  dplyr::select(starts_with("t")) %>% 
  ff_glimpse()
```
We have duplicated hmdb (positive and negative)
```{r}
known_metabolites %>%
  separate(row_id, into = c("charge", "id")) %>% 
  mutate(id = as.character(id)) %>% 
  ggplot(aes(x=charge, y=id, fill= charge))+
  geom_tile()+
  theme_classic()
```

```{r}
known_metabolites %>%
  separate(row_id, into = c("charge", "id")) %>% 
  filter(duplicated(id)) %>% 
  full_join(., omics_data %>%
              separate(row_id, into = c("charge", "id")) %>% 
              filter(!duplicated(id)),
            by="id") %>% 
  dplyr::select(charge.x, id, charge.y) %>% 
  filter(is.na(charge.x) | is.na(charge.y)) %>% 
  pivot_longer(cols = c(charge.x, charge.y)) %>% 
  filter(!is.na(value)) %>% 
  ggplot(aes(x=value, fill= value))+
  geom_bar()+
  theme_classic()
```
The data is imputed with minimum of the expression after selecting for negative metabolites.

```{r data prep}
mldata <- mldata %>% 
  # filter metabolites in positive or negative detection
  filter(str_detect(row_id, "neg")) %>% 
  # Imputation samples data
  mutate(across((starts_with("t")), .fns = ~ replace_na(., min(., na.rm = TRUE)))) %>% 
  mutate_if(is.character, factor) %>% 
  dplyr::select(hmdb, row_m_z, row_retention_time, 
                starts_with("t"))
```

